Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-05-27 doc/LinksToDAG.tex
added note that there are 122 link types

It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.  However, this is not quite true.  Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.  It produces undirected connected planar graphs.  The annotation conventions are implicit in the detailed Link Grammar lexicon,%
\footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.  The link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.} 
which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.  As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.  For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.

@@

Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.  If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.  In this way, the empirical distribution of sentences guides our assignment of directions.

>>>>

It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.  However, this is not quite true.  Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.  It produces undirected connected planar graphs.  The annotation conventions are implicit in the detailed Link Grammar lexicon,%
\footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.  The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.} 
which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.  As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.  For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.

@@

Thus, we chose to use a sample of {\em actual} sentences parsed by the link grammar, and to seek a direction mapping so that {\em these} parses would not violate our constraints after directions were imposed.  If no such mapping exists, then we are willing to orient a few edge tokens in the wrong direction to ensure that the parses are still well-formed---but we minimize the number of such violations.  In this way, the empirical distribution of sentences guides our assignment of directions.  The resulting directed corpus can be used for research on multi-headed dependency parsing.
